%Formatting
\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{layout}
\usepackage{setspace}


\bibliographystyle{apalike}
\doublespacing
\topmargin -72pt
\oddsidemargin 0pt
\textheight 10in
\textwidth 455pt
\headsep 36pt

\itemsep -10pt

\renewcommand{\thepage}{Page~\arabic{page}}

\begin{document}
%\layout

\title{All the World: Scaffolding}
\author{William Jardee}
\maketitle

\textbf{Significant World Building Facets:}
\begin{enumerate}
\item Magic in this world is a well-established, unexplained phenomenon that seems to ``pour out of one's life essence," requiring attunement with the physical reality of one's surroundings and the exertion of energy, an amount proportional to the impact of the spell being cast. 
\item The development of science is slightly delayed because individuals have relied on magic instead of the invention.
\item Regulations, enforced by a governmental body, have been put in place to allow for the use of ``small" magic and control ``big" magic.
\item Current break-throughs in theoretical physics (Weavon-theory) have introduced a physical explanation for magic. 
\item Both governmental bodies and individual companies want this information for themselves, so there is a mad dash to get exclusive control over this information.
\end{enumerate}

\textbf{Examples of Challenges Characters May Face:}
\begin{enumerate}
\item What are the contractual, moral, and personal obligations with this information? How does aligning with one introduce risk from the others?
\item What should they do with the {\sl theoretical} break-through? Do they need to do experimental work before sharing it? What if other people publish their work before they can get experimental verification? 
\end{enumerate}

\end{document}