%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Graduate Curriculum Vitae
% LaTeX Template
% Version 1.2 (3/28/15)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute 
% (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
% Modified by:
% Daniel L Marks <xleafr@gmail.com> 3/28/2015
%
% Important note:
% This template requires the res.cls file to be in the same directory as the
% .tex file. The res.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%-------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% You can have multiple style options the legal options ones are:
%
%   centered:	the name and address are centered at the top of the page 
%				(default)
%
%   line:		the name is the left with a horizontal line then the address to
%				the right
%
%   overlapped:	the section titles overlap the body text (default)
%
%   margin:		the section titles are to the left of the body text
%		
%   11pt:		use 11 point fonts instead of 10 point fonts
%
%   12pt:		use 12 point fonts instead of 10 point fonts
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[margin]{res}  

% Default font is the helvetica postscript font
\usepackage{helvet}

% Increase text height
\textheight=700pt

\begin{document}

%-------------------------------------------------------------------------------
%	NAME AND ADDRESS SECTION
%-------------------------------------------------------------------------------
\name{William V. Jardee}

% Note that addresses can be used for other contact information:
% -phone numbers
% -email addresses
% -linked-in profile

\address{2411 Wheeler Dr. Unit D\\Bozeman, MT 59715\\United States of America}
\address{willjardee@gmail.com\\(406) 836-2338}

% Uncomment to add a third address
%\address{Address 3 line 1\\Address 3 line 2\\Address 3 line 3}
%-------------------------------------------------------------------------------

\begin{resume}

%-------------------------------------------------------------------------------
%	EDUCATION SECTION
%-------------------------------------------------------------------------------
\section{EDUCATION}
\textbf{Montana State University}, Bozeman, MT\\
{\sl B.S., Summa Cum Laude} (Phi Kappa Phi Honors Society), Professional Physics \\
{\sl Minors in}, Computer Science and Mathematics, May 2022\hfill GPA: 3.94/4.0\\


%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
%	PROJECTS SECTION
%-------------------------------------------------------------------------------
\section{PROJECTS}
\par
\textbf{Relativistic Runaway Electrons and Lightning Discharge; A Qualitative Overview}: 
Building blocks of the RREA Theory, alongside motivations, computational and experimental evidence. Survey of step leaders and the related TGF emissions.
\par
\textbf{RREA Propagation Theory; A Theoretical and Computational Overview}:
A delve into the theoredical derivation of RREA theory and the implementation of complete Monte Carlo simulations of particle propegation in stormclouds. \textit{proposed project}
\par 
\textbf{Introduction to Computational Physics}
A collection of Python, \LaTeX , and essential tools, like Jupyter Notebook, overviews along with more detailed delves into important topics like ODE/PDE solvers, Monte Carlo simulations, and large data set analysis. Including asides about AI, ML, TDA, computational theory, and modern topics like dynamical systems. \textit{ongoing project}

%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
%	COMPUTER SKILLS SECTION
%-------------------------------------------------------------------------------
\section{TECHNICAL\\SKILLS}

\textbf{Languages}: Python, GitHub, Java, C/C++, \LaTeX, HTML, CSS, Matlab, Mathematica, Excel
\\
\textbf{Mathematics}: Linear Algebra, Dynamical/Chaotic Systems, Computation Theory 
\textbf{Physics}: Particle Physics, Observational Astronomy, General Relativity

%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
%	EXPERIENCE SECTION
%-------------------------------------------------------------------------------
% Modify the format of each position
\begin{format}
\title{l}\employer{r}\\
\dates{l}\location{r}\\
\body\\
\end{format}
%-------------------------------------------------------------------------------

\section{TEACHING EXPERIENCE}
\employer{Allen Yarnell Student Success Center}
\location{Montana State University, Bozeman}
\dates{July 2021 - Present}
\title{\textbf{Hillman's Scholars Tutor}}
\begin{position}
Educating underprivileged college students in introductory math, physics, computer science, and humanities courses.
\end{position}
\employer{Mathematics Department}
\location{Montana State University, Bozeman}
\dates{Aug 2021 - Present}
\title{\textbf{Math Stats Center Tutor}}
\begin{position}
Guide students to discover their own answers and understanding in classes ranging from introductory algebra to differential equations.
\end{position}
\employer{Physics Department}
\location{Montana State University, Bozeman}
\dates{Jan 2021 - May 2021}
\title{\textbf{Proctor/Grader (PHSX 207)}}
\begin{position}
Grading weekly homework and exams of algebra based introductory physics course. 
\end{position}
\employer{Physic Department}
\location{Montana State University, Bozeman}
\dates{Aug 2020 - Nov 2020}
\title{\textbf{Student Lab Assistant (PHSX 205)}}
\begin{position}
Guided students of the introductory physics course through kinematic labs during a weekly lab. 
\end{position}
\employer{Allen Yarnell Student Success Center}
\location{Montana State University, Bozeman}
\dates{Aug 2019 - May 2020}
\title{\textbf{Smarty Cats Tutor}}
\begin{position}
Made and personalized appointments with students to cover essential STEM topics.
\end{position}
\employer{The Rock Youth Center}
\location{Montana State University, Bozeman}
\dates{Oct 2019 - March 2020}
\title{\textbf{Volunteer Stem Tutor}}
\begin{position}
Held open tutoring hours for high school students who struggles in STEM subjects
\end{position}


%-------------------------------------------------------------------------------

\section{RESEARCH EXPERIENCE}
\employer{Dr. John Sample's Lab}
\location{Montana State University, Bozeman}
\dates{Aug 2020 - Dec 2020}
\title{\textbf{Undergraduate Researcher}}
\begin{position}
Analyzed the performance of a soft x-ray spectrometer to be attached to the IMPRESS CUBE-SAT. 
\end{position}
\employer{Dr. Rufus Cone's Lab}
\location{Montana State University, Bozeman}
\dates{Jan 2020 - Apr 2020}
\title{\textbf{Undergraduate Researcher}}
\begin{position}
Studied the theory behind, and attempted to use, an ellipsometer to measure the thickness of thin wafers.
\end{position}

%-------------------------------------------------------------------------------

\section{MISC EXPERIENCE}
\employer{Society of Physics Students at Montana State University}
\location{Montana State University, Bozeman}
\dates{Feb 2020 - Jan 2022}
\title{\textbf{Club Treasurer}}
\begin{position}
Handled club finance and lead many efforts in stirring interest in science communication and computational physics. 
\end{position}

%-------------------------------------------------------------------------------

\section{AWARDS/ GRANTS}
\textbf{Physics Departmental Scholarship}
\begin{itemize}
\item \textbf{Norman Mac Rugheimer Scholarship} \hfill Aug 2021
\item \textbf{Asbridge Physics Scholarship} \hfill Aug 2020
\end{itemize}
\textbf{Montana University Systems Scholarship}\hfill\hfil May 2018

%-------------------------------------------------------------------------------

\section{POSTERS/ PRESENTATIONS}
\textbf{RREA Propagation Theory} \hfill Oct 2021\\
\textit{SPS Undergraduate Colloquium}
	
%-------------------------------------------------------------------------------

\section{OUTREACH}
\textbf{Grossology} \hfill Oct 2021\\
\textit{Museum of the Rockies}\\
\textbf{Liquid Nitrogen Ice Cream} \hfill Oct 2021\\
\textit{Society of Physics Students}\\
\textbf{Careers in Industry Panel; Moderator} \hfill Oct 2020, Mar 2021\\
\textit{Society of Physics Students}


%-------------------------------------------------------------------------------
%	Interests
%-------------------------------------------------------------------------------
\section{INTERESTS}
Science communication, computational physics, particle physics, chaotic systems, macroscopic weather systems, RREA propagation theory
%-------------------------------------------------------------------------------
\end{resume}
\end{document}